---
title: "122 Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Import datasets into R
```{r}
winequality.red <- read.csv("C:/Users/rtora/Downloads/winequality-red.csv", sep=";")
View(winequality.red)
```

```{r}
winequality.white <- read.csv("C:/Users/rtora/Downloads/winequality-white.csv", sep=";")
View(winequality.white)
```
#Add a new column to the datasets to id red and white
#white = 0  red = 1
```{r}
winequality.white$id = 0
winequality.red$id = 1

```
#combine datasets into 1
```{r}
winequality = rbind(winequality.red,winequality.white)
View(winequality)
```

```{r}
attach(winequality)
plot(quality[id == 1], alcohol[id == 1], col = "black", main = "quality vs alcohol content",ylab="alcohol",xlab="quality")
points(quality[id==0],alcohol[id==0],col = "yellow")
```
#create a scatterplot matrix and determine correlations between predictors
```{r}
full_model = lm(formula = quality ~ fixed.acidity +volatile.acidity  + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol + id)
pairs(winequality)
cor(subset(winequality, select = -c(quality)))
```
#create a step function to determine an appropriate linear model

```{r}
library("car", lib.loc="~/R/win-library/3.6")
vif(full_model)
```
#we can drop density from the predictors because it has the largest VIF
#lets do it agian with the new predictors
```{r}
vif(lm(formula = quality ~ fixed.acidity +volatile.acidity  + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide  + pH + sulphates + alcohol + id))
```
#we can drop id from predictors because it is over 5
#let's do it again
```{r}
vif(lm(formula = quality ~ fixed.acidity +volatile.acidity  + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide  + pH + sulphates + alcohol))
```
#finally we can remove total.sulfur.dioxide because of its higher correlation with free.sulfur dioxide and its larger vif value
```{r}
vif(lm(formula = quality ~ fixed.acidity +volatile.acidity  + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide  + pH + sulphates + alcohol ))
```
#since everything is smaller than 2 we can keep the rest of the predictors
#fit full model
```{r}
full_model = lm(formula = quality ~ fixed.acidity +volatile.acidity  + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide  + pH + sulphates + alcohol )

```
#diagnostic plots for the full model
```{r}
plot(full_model)
plot(winequality$fixed.acidity, resid(full_model), main = "Residuals vs fixed acidity", ylab = "fixed acidity",xlab= "residuals")
plot(winequality$volatile.acidity, resid(full_model), main = "Residuals vs volatile acidity", ylab = "volatile acidity",xlab= "residuals")
plot(winequality$citric.acid, resid(full_model), main = "Residuals vs citric acid", ylab = "citric acid",xlab= "residuals")
plot
plot(winequality$chlorides, resid(full_model), main = "Residuals vs chlorides", ylab = "chlorides",xlab= "residuals")
plot(winequality$free.sulfur.dioxide, resid(full_model), main = "Residuals vs free sulfur dioxide", ylab = "free sulfur dioxide",xlab= "residuals")
plot(winequality$pH, resid(full_model), main = "Residuals vs pH", ylab = "pH",xlab= "residuals")
plot(winequality$sulphates, resid(full_model), main = "Residuals vs sulphates", ylab = "sulphates",xlab= "residuals")
plot(winequality$alcohol, resid(full_model), main = "Residuals vs alcohol", ylab = "alcohol",xlab= "residuals")
```
#The residual plot has a horizontal band around the 0 line, with no discernable pattern in the plot. Thus we satisfy the Linearity and Equal Variane LINE conditions.
#stepwise regression
#use log of free sulfur dioxide chlorides and sulphates to normalize the data

```{r}
library(Rcmdr)
mod0 = lm(quality~.)
mod.upper = lm(formula = quality ~ fixed.acidity +volatile.acidity  + citric.acid + residual.sugar + log(chlorides) + log(free.sulfur.dioxide)  + pH + log(sulphates) + alcohol )
stepmodel = stepwise(mod0,scope=list(lower = mod0, upper = mod.upper))
```
```{r}
anova(stepmodel)
summary(stepmodel)
```

#since all of the values seem to have large F-values we will keep them
```{r}
library(olsrr)
cooksd <- cooks.distance(mod.upper)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
```
#testing the model without the influential points
```{r}
winequality.no.outlier = winequality[-c(653),]
mod0 = lm(quality~1,data = winequality.no.outlier)
mod.upper = lm(formula = quality ~ fixed.acidity +volatile.acidity  + citric.acid + residual.sugar + log(chlorides) + log(free.sulfur.dioxide)  + pH + log(sulphates) + alcohol , data = winequality.no.outlier)
stepmodel = step(mod0,scope=list(lower = mod0, upper = mod.upper))
```

```{r}
cooksd <- cooks.distance(mod.upper)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
```

```{r}
summary(mod.upper)
```

#testing the regression model
```{r}
set.seed(27)
rows = sample(nrow(winequality))
shuffle_winequality = winequality[rows, ]
testing = shuffle_winequality[1:1000,]
```
#test the number correct
```{r}
correct = 0
total = 1000
for(i in 1:1000){
 if(testing[i,]$quality == round(1.331817 + (0.354809*(testing[i,]$alcohol)) + (-1.365576*testing[i,]$volatile.acidity) + (log(testing[i,]$sulphates)     *0.382366) + (log(testing[i,]$free.sulfur.dioxide)*0.124498 )+(testing[i,]$residual.sugar * 0.017083 )+(testing[i,]$pH*0.237243)+(testing[i,]$fixed.acidity*0.042343)+(testing[i,]$citric.acid * -0.250106) )){
   correct = correct + 1
}
}
testing[1,]$alcohol
print(correct)
print(correct/total)
```
#Therefore our model has about 54% accuracy which is much more accurate than a random guess (10%)
```{r}
correct = 0
total = 1000
for(i in 1:1000){
 if(testing[i,]$quality == round(1.234678 + (0.354598*(testing[i,]$alcohol)) + (-1.354607*testing[i,]$volatile.acidity) + (log(testing[i,]$sulphates)     *0.388967) + (log(testing[i,]$free.sulfur.dioxide)*0.125422 )+(testing[i,]$residual.sugar * 0.017203 )+(testing[i,]$pH*0.243014)+(testing[i,]$fixed.acidity*0.046388)+(testing[i,]$citric.acid * -0.250998) )){
   correct = correct + 1
}
}
testing[1,]$alcohol
print(correct)
print(correct/total)
```


```{r}
library(corrgram)
corrgram(winequality, lower.panel=panel.shade, upper.panel=panel.ellipse)
```

```{r}
plot(mod.upper)
plot(winequality.no.outlier$fixed.acidity, resid(mod.upper), main = "Residuals vs fixed acidity", ylab = "fixed acidity",xlab= "residuals")
plot(winequality.no.outlier$volatile.acidity, resid(mod.upper), main = "Residuals vs volatile acidity", ylab = "volatile acidity",xlab= "residuals")
plot(winequality.no.outlier$citric.acid, resid(mod.upper), main = "Residuals vs citric acid", ylab = "citric acid",xlab= "residuals")
plot
plot(winequality.no.outlier$chlorides, resid(mod.upper), main = "Residuals vs chlorides", ylab = "chlorides",xlab= "residuals")
plot(winequality.no.outlier$free.sulfur.dioxide, resid(mod.upper), main = "Residuals vs free sulfur dioxide", ylab = "free sulfur dioxide",xlab= "residuals")
plot(winequality.no.outlier$pH, resid(mod.upper), main = "Residuals vs pH", ylab = "pH",xlab= "residuals")
plot(winequality.no.outlier$sulphates, resid(mod.upper), main = "Residuals vs sulphates", ylab = "sulphates",xlab= "residuals")
plot(winequality.no.outlier$alcohol, resid(mod.upper), main = "Residuals vs alcohol", ylab = "alcohol",xlab= "residuals")
```
```{r}
anova(mod.upper)
```
```{r}
mod.upper2 = lm(formula = quality ~ fixed.acidity +volatile.acidity + residual.sugar + log(chlorides) + log(free.sulfur.dioxide)  + pH + log(sulphates) + alcohol , data = winequality.no.outlier)
anova(mod.upper2)
plot(mod.upper2)
```
```{r}
res<-t.test(winequality.red$quality,winequality.white$quality)
res
```

